---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
unzip_activity <- unzip("./activity.zip")
act <- read.csv(unzip_activity)
head(act)
tail(act)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(dplyr)
day_act <- group_by(act, date)
step_perday_act <- summarize(day_act, sums = sum(steps), means=mean(steps), medians=median(steps))
hist(step_perday_act$sums)
step_perday_act$means 
step_perday_act$medians
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
inte_act <- group_by(act, interval)
step_inte_act <- summarize(inte_act, means = mean(steps, na.rm=TRUE))
plot(step_inte_act$means, type="l")
step_inte_act[which(step_inte_act$means==max(step_inte_act$means)),]
```

## Imputing missing values
```{r, echo=TRUE}
sum(is.na(act$steps))
```
### Devise a strategy for filling mean of that day in all of the missing values in the dataset
```{r, echo=TRUE}
new_act <- act
act_row <- nrow(act)
for (i in 1:act_row){
  if(is.na(new_act[i, ]$steps)){
    na_date <- new_act[i, ]$date
    new_act[i, ]$steps <- step_perday_act[step_perday_act$date ==na_date, ]$means
  }
}
```

### mean and median of the new_act dateframe
```{r, echo=TRUE}
library(dplyr)
day_newact <- group_by(new_act, date)
step_perday_newact <- summarize(day_newact, sums = sum(steps), means=mean(steps), medians=median(steps))
hist(step_perday_newact$sums)
step_perday_newact$means 
step_perday_newact$medians
```

### why new_act is not different from act
```{r, echo=TRUE}
table(act[is.na(act$steps), ]$date)
```

#### on the day when there is any NA data, there is no other data then NA. So, the new_act datafrem didn't change
### So, I changed the way of filling NA data. I devised a strategy for filling mean of the 5-minutes interval in all of the missing values in the dataset
```{r, echo=TRUE}
new_act <- act
act_row <- nrow(act)
for (i in 1:act_row){
  if(is.na(new_act[i, ]$steps)){
    na_inte <- new_act[i, ]$interval
    new_act[i, ]$steps <- step_inte_act[step_inte_act$interval ==na_inte, ]$means
  }
}
```

### mean and median of the new_act dateframe
```{r, echo=TRUE}
library(dplyr)
day_newact <- group_by(new_act, date)
step_perday_newact <- summarize(day_newact, sums = sum(steps), means=mean(steps), medians=median(steps))
hist(step_perday_newact$sums)
step_perday_newact$means 
step_perday_newact$medians
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
library(lubridate)
library(ggplot2)
library(dplyr)
weekday_vec <- c("Mon","Tue","Wed","Thu","Fri")
weekend_vec <- c("Sat", "Sun")
weekday <- data.frame()
weekend <- data.frame()

for (i in 1:nrow(new_act)){
  if (wday(new_act[i,]$date,label=TRUE,locale = "en_US.UTF-8") %in% weekday_vec){
    weekday <- bind_rows(weekday, new_act[i,])
  }
  else{
    weekend <- bind_rows(weekend, new_act[i,])
  }
}

inte_weekend <- group_by(weekend, interval)
step_inte_weekend <- summarize(inte_weekend, means = mean(steps))
step_inte_weekend$wday <- "weekend"
inte_weekday <- group_by(weekday, interval)
step_inte_weekday <- summarize(inte_weekday, means = mean(steps))
step_inte_weekday$wday <- "weekday"

final_act <- bind_rows(step_inte_weekend, step_inte_weekday)
p_0 <- ggplot(final_act, aes(x = interval, y = means)) +geom_line()+facet_grid(rows = vars(wday))+ theme_bw() 
p_0 +labs(x = "Iinterval", y = "Number of steps")
```
